---
title: "Graphseg for segmentation of spatial data"
author: "Vivien Goepp"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
  ---
  
  ```{r setup, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
```
```{r}
library(sf) # manipulate geometrical data
library(magrittr) # use pipe operator
library(dplyr) # for dataframe manipulation
library(Matrix) # use sparse matrices
library(igraph) # manipulate graphs
library(graphseg)
```

# Loading the data

```{r}
data("utrecht_district")
data("graph_utrecht_district")

plot(utrecht_district["geometry"])
```

# Generate noisy data
Define a piece-wise constant signal on the adjacency graph, and define a noisy signal thereof:
```{r}
baseline <- data.frame("municip_code" = levels(utrecht_district$municip_code),
                       "pc_signal" = rnorm(nlevels(utrecht_district$municip_code), 0, 10))
utrecht_district <- utrecht_district %>% 
  left_join(baseline, by = "municip_code") # define piece-wise constant signal
sigma_sq <- 1
utrecht_district$noisy_signal <- utrecht_district$pc_signal + 
  rnorm(nrow(utrecht_district), 0, sigma_sq) # define noisy signal
plot(utrecht_district["noisy_signal"])
```

# Fused regularized signal approximation
Run `agraph`:
```{r}
lambda <- 10 ^ seq(-3, 3, length = 50) # define a sequence of penalties
res <- agraph(utrecht_district$noisy_signal, graph_utrecht_district, lambda)
```
Plot regularization path:
```{r}
matplot(log10(lambda), res$result, type = "l")
abline(v = log10(lambda)[sapply(res[c("aic", "bic", "gcv")], which.min)],
       col = "red")
```
Plot the estimated piecewise constant signal:
```{r}
utrecht_district$seg_aic <- res$result[which.min(res$aic), ]
utrecht_district$seg_bic <- res$result[which.min(res$bic), ]
plot(utrecht_district["seg_aic"])
plot(utrecht_district["seg_bic"])

utrecht_district_fused <- utrecht_district %>% 
  dplyr::mutate(group = cut(seg_bic, breaks = sort(unique(seg_bic)),
                            include.lowest = TRUE)) %>%
  dplyr::group_by(group) %>%
  dplyr::summarize(seg_bic = mean(seg_bic), do_union = TRUE)

plot(utrecht_district_fused["seg_bic"])
```

Display the fused estimation as a one-dimensional signal:
```{r}
plot(utrecht_district$baseline)
points(res$result[which.min(res$bic), ], pch = 3, col = "red")
```


